---
title: "Exploratory Data Analysis (Coursera)"
output: html_notebook
---
# Exploratory Graphs (1)
```{r}
pollution <- read.csv("avgpm25.csv", colClasses = c("numeric", "character", "factor", "numeric", "numeric"))
head(pollution)
```
## One dimension (Summaries of Data) 
* Five-number summary 
* Boxplot
* Histogram
* Density Plot
* Bar plot

```{r}
summary(pollution$pm25)
```

```{r}
boxplot(pollution$pm25, col = "blue")
abline(h = 12)
```

```{r}
hist(pollution$pm25, col = "green", breaks = 100)
rug(pollution$pm25)
abline(v = 12, lwd = 2)
abline(v = median(pollution$pm25), col = "magenta", lwd = 4)
```

```{r}
barplot(table(pollution$region), col = "wheat", main = "Number of Counties in Each Region")
```

## Two dimension 
* Multiple/Overlayed 1-D Plots (Lattice/ggplot2)
* Scatterplots
* Smooth scatterplots
> 2 dimensions: 
* Overlayed/multiple 2-D plots; coplots
* Use color, size, shape to add dimensions
* Spinning plots
* Actual 3-D plots (not that useful)

Multiple Box Plots
```{r}
boxplot(pm25 ~ region, data = pollution, col = c("red", "blue"))
```

```{r}
par(mfrow = c(2,1), mar = c(4,4,2,1))
hist(subset(pollution, region == 'east')$pm25, col = "green")
hist(subset(pollution, region == 'west')$pm25, col = "green")

```
Scatterplot
```{r}
with(pollution, plot(latitude, pm25))
abline(h = 12, lwd = 2, lty = 2)

with(pollution, plot(latitude, pm25, col = region))
abline(h = 12, lwd = 2, lty = 2)
```


```{r}
par(mfrow = c(1,2), mar = c(5,4,2,1))
with(subset(pollution, region == "west"), plot(latitude, pm25, main = "West"))
with(subset(pollution, region == "east"), plot(latitude, pm25, main = "East"))

```

# Plotting Systems in R
THREE OPTIONS
1. Base Plotting System => "Artist's Palette" model 
- usually starts with plot function
- convenient and easy 
- can't go back once a plot has started
- difficult to 'translate' to others
```{r}
library(datasets)
data(cars)
with(cars, plot(speed, dist))
```
2. Lattice System 
- plots are created with single function call (xyplot, bwplot, etc...)
- Most useful for conditioning types of plots: looking at how y changes with x across levels of z
- Automates margins, structure, etc... 
- Sometimes awkward to specify an entire plot in a single function call


3. ggplot2 System 
- splits the difference between base and lattice in a number of ways
- automatically deals with misc stuff
- superficial similarity to lattice but generally easier to use

```{r}
library("ggplot2")
library(datasets)
str(mpg)
qplot(displ, hwy, data = mpg)

qplot(displ, hwy, color = drv, data = mpg)
qplot(displ, data = mpg)
```


## Base Plotting System

## The Process of Making a Plot

When making a plot one must first make a few considerations (not
necessarily in this order):

- Where will the plot be made? On the screen? In a file? 

<!--
  - The default in Unix is `x11`; on Windows it is `windows`; on Mac
    OS X it is `quartz`
-->

- How will the plot be used?
  - Is the plot for viewing temporarily on the screen?  
  - Will it be presented in a web browser?
  - Will it eventually end up in a paper that might be printed? 
  - Are you using it in a presentation?

- Is there a large amount of data going into the plot? Or is it just a
  few points?

- Do you need to be able to dynamically resize the graphic?

---

## The Process of Making a Plot

- What graphics system will you use: base, lattice, or ggplot2? These
  generally cannot be mixed.

- Base graphics are usually constructed piecemeal, with each aspect of
  the plot handled separately through a series of function calls; this
  is often conceptually simpler and allows plotting to mirror the
  thought process

- Lattice graphics are usually created in a single function call, so
  all of the graphics parameters have to specified at once; specifying
  everything at once allows R to automatically calculate the necessary
  spacings and font sizes.

- ggplot2 combines concepts from both base and lattice graphics but
  uses an independent implementation

We focus on using the **base plotting system** to create graphics on
the **screen device**.

---

## Base Graphics

Base graphics are used most commonly and are a very powerful system for creating 2-D graphics.

- There are two *phases* to creating a base plot
  - Initializing a new plot
  - Annotating (adding to) an existing plot

- Calling `plot(x, y)` or `hist(x)` will launch a graphics device (if
  one is not already open) and draw a new plot on the device

- If the arguments to `plot` are not of some special class, then the
  _default_ method for `plot` is called; this function has _many_
  arguments, letting you set the title, x axis label, y axis label,
  etc.

- The base graphics system has _many_ parameters that can set and
  tweaked; these parameters are documented in `?par`; it wouldn’t hurt
  to try to memorize this help page!

---

## Simple Base Graphics: Histogram

```{r,fig.height=5}
library(datasets)
hist(airquality$Ozone)  ## Draw a new plot
```

---

## Simple Base Graphics: Scatterplot

```{r,fig.height=5}
library(datasets)
with(airquality, plot(Wind, Ozone))
```

---

## Simple Base Graphics: Boxplot

```{r}
library(datasets)
airquality <- transform(airquality, Month = factor(Month))
boxplot(Ozone ~ Month, airquality, xlab = "Month", ylab = "Ozone (ppb)")

```


## Some Important Base Graphics Parameters

Many base plotting functions share a set of parameters. Here are a few
key ones:

- `pch`: the plotting symbol (default is open circle) - plotting character

- `lty`: the line type (default is solid line), can be dashed, dotted, etc.

- `lwd`: the line width, specified as an integer multiple

- `col`: the plotting color, specified as a number, string, or hex
  code; the `colors()` function gives you a vector of colors by name

- `xlab`: character string for the x-axis label

- `ylab`: character string for the y-axis label


---

## Some Important Base Graphics Parameters

The `par()` function is used to specify *global* graphics parameters
that affect all plots in an R session. These parameters can be
overridden when specified as arguments to specific plotting functions.

- `las`: the orientation of the axis labels on the plot
- `bg`: the background color
- `mar`: the margin size
- `oma`: the outer margin size (default is 0 for all sides)
- `mfrow`: number of plots per row, column (plots are filled row-wise) 
- `mfcol`: number of plots per row, column (plots are filled column-wise)

---

## Some Important Base Graphics Parameters

Default values for global graphics parameters

---

## Some Important Base Graphics Parameters

Default values for global graphics parameters

```{r}
par("bg")
par("mar")
par("mfrow")
```

## Base Plotting Functions

- `plot`: make a scatterplot, or other type of plot depending on the
  class of the object being plotted

- `lines`: add lines to a plot, given a vector x values and a
  corresponding vector of y values (or a 2-column matrix); this
  function just connects the dots

- `points`: add points to a plot
- `text`: add text labels to a plot using specified x, y coordinates
- `title`: add annotations to x, y axis labels, title, subtitle, outer margin 
- `mtext`: add arbitrary text to the margins (inner or outer) of the plot 
- `axis`: adding axis ticks/labels


## Base Plot with Annotation

```{r,fig.height=5}
library(datasets)
with(airquality, plot(Wind, Ozone))
title(main = "Ozone and Wind in New York City")  ## Add a title
```


## Base Plot with Annotation

```{r,fig.height=5}
with(airquality, plot(Wind, Ozone, main = "Ozone and Wind in New York City"))
with(subset(airquality, Month == 5), points(Wind, Ozone, col = "blue"))
```


## Base Plot with Annotation

```{r,fig.height=5}
with(airquality, plot(Wind, Ozone, main = "Ozone and Wind in New York City", type = "n")) # n is invisible
with(subset(airquality, Month == 5), points(Wind, Ozone, col = "blue"))
with(subset(airquality, Month != 5), points(Wind, Ozone, col = "red"))
legend("topright", pch = 1, col = c("blue", "red"), legend = c("May", "Other Months"))
```


## Base Plot with Regression Line

```{r,fig.height=5}
with(airquality, plot(Wind, Ozone, main = "Ozone and Wind in New York City", pch = 20))
model <- lm(Ozone ~ Wind, airquality)
abline(model, lwd = 2)
```



## Multiple Base Plots

```{r,fig.height=5,fig.width=14}
par(mfrow = c(1, 2))
with(airquality, {
	plot(Wind, Ozone, main = "Ozone and Wind")
	plot(Solar.R, Ozone, main = "Ozone and Solar Radiation")
})
```



## Multiple Base Plots

```{r,fig.height=4,fig.width=12}
par(mfrow = c(1, 3), mar = c(4, 4, 2, 1), oma = c(0, 0, 2, 0)) # mar => margin ; oma => outer margin
with(airquality, {
	plot(Wind, Ozone, main = "Ozone and Wind")
	plot(Solar.R, Ozone, main = "Ozone and Solar Radiation")
	plot(Temp, Ozone, main = "Ozone and Temperature")
	mtext("Ozone and Weather in New York City", outer = TRUE)
})
```

## Summary

* Plots in the base plotting system are created by calling successive
  R functions to "build up" a plot

* Plotting occurs in two stages:
  - Creation of a plot
  - Annotation of a plot (adding lines, points, text, legends)

* The base plotting system is very flexible and offers a high degree
  of control over plotting


Ex) 
```{r}
x <- rnorm(100) 
hist(x)
y <- rnorm(100)
par(mar = c(4,4,2,2))
plot(x,y, pch = 18)
title("Scatterplot")
legend("topleft", legend = "Data", pch=18)
fit <- lm(y~x)
abline(fit)
```

    
```{r}
example(points)
```



## The Lattice Plotting System

The lattice plotting system is implemented using the following packages:

- *lattice*: contains code for producing Trellis graphics, which are
   independent of the “base” graphics system; includes functions like
   `xyplot`, `bwplot`, `levelplot`

- *grid*: implements a different graphing system independent of the
   “base” system; the *lattice* package builds on top of *grid*
   - We seldom call functions from the *grid* package directly

- The lattice plotting system does not have a "two-phase" aspect with
  separate plotting and annotation like in base plotting

- All plotting/annotation is done at once with a single function call



## Lattice Functions

- `xyplot`: this is the main function for creating scatterplots 
- `bwplot`: box-and-whiskers plots (“boxplots”)
- `histogram`: histograms
- `stripplot`: like a boxplot but with actual points 
- `dotplot`: plot dots on "violin strings"
- `splom`: scatterplot matrix; like `pairs` in base plotting system 
- `levelplot`, `contourplot`: for plotting "image" data


## Lattice Functions

Lattice functions generally take a formula for their first argument, usually of the form

```r
xyplot(y ~ x | f * g, data)
```

- We use the *formula notation* here, hence the `~`.

- On the left of the ~ is the y-axis variable, on the right is the
  x-axis variable

- f and g are _conditioning variables_ — they are optional
  - the * indicates an interaction between two variables

- The second argument is the data frame or list from which the
  variables in the formula should be looked up

  - If no data frame or list is passed, then the parent frame is used.

- If no other arguments are passed, there are defaults that can be used.

---
## Simple Lattice Plot

```{r,fig.height=5}
library(lattice)
library(datasets)
## Simple scatterplot
xyplot(Ozone ~ Wind, data = airquality)
```



## Simple Lattice Plot

```{r,fig.width=12,fig.height=4}
library(datasets)
library(lattice)
## Convert 'Month' to a factor variable
airquality <- transform(airquality, Month = factor(Month)) 
xyplot(Ozone ~ Wind | Month, data = airquality, layout = c(5, 1))
```

## Lattice Behavior

Lattice functions behave differently from base graphics functions in
one critical way.

- Base graphics functions plot data directly to the graphics device
  (screen, PDF file, etc.)

- Lattice graphics functions return an object of class **trellis**

- The print methods for lattice functions actually do the work of
  plotting the data on the graphics device.

- Lattice functions return "plot objects" that can, in principle, be
  stored (but it’s usually better to just save the code + data).

- On the command line, trellis objects are *auto-printed* so that it
  appears the function is plotting the data



## Lattice Behavior

```{r,fig.height=4,fig.width=5}
p <- xyplot(Ozone ~ Wind, data = airquality)  ## Nothing happens!
print(p)  ## Plot appears
```
```{r,eval=FALSE}
xyplot(Ozone ~ Wind, data = airquality)  ## Auto-printing
```

## Lattice Panel Functions

* Lattice functions have a **panel function** which controls what
  happens inside each panel of the plot.

* The *lattice* package comes with default panel functions, but you
  can supply your own if you want to customize what happens in each
  panel

* Panel functions receive the x/y coordinates of the data points
  in their panel (along with any optional arguments)




## Lattice Panel Functions

```{r,fig.height=4,fig.width=8}
set.seed(10)
x <- rnorm(100)
f <- rep(0:1, each = 50)
y <- x + f - f * x+ rnorm(100, sd = 0.5)
f <- factor(f, labels = c("Group 1", "Group 2"))
xyplot(y ~ x | f, layout = c(2, 1))  ## Plot with 2 panels
```



## Lattice Panel Functions

```{r,fig.height=4}
## Custom panel function
xyplot(y ~ x | f, panel = function(x, y, ...) {
       panel.xyplot(x, y, ...)  ## First call the default panel function for 'xyplot'
       panel.abline(h = median(y), lty = 2)  ## Add a horizontal line at the median
})
```



## Lattice Panel Functions: Regression line

```{r,fig.height=4}
## Custom panel function
xyplot(y ~ x | f, panel = function(x, y, ...) {
               panel.xyplot(x, y, ...)  ## First call default panel function
               panel.lmline(x, y, col = 2)  ## Overlay a simple linear regression line
       })
```

## Summary

* Lattice plots are constructed with a single function call to a core
  lattice function (e.g. `xyplot`)

* Aspects like margins and spacing are automatically handled and
  defaults are usually sufficient

* The lattice system is ideal for creating conditioning plots where
  you examine the same kind of plot under many different conditions

* Panel functions can be specified/customized to modify what is
  plotted in each of the plot panels


# GGPLOT2


## What is ggplot2?

- An implementation of _The Grammar of Graphics_ by Leland Wilkinson
- Written by Hadley Wickham (while he was a graduate student at Iowa State)
- A “third” graphics system for R (along with __base__ and __lattice__)
- Available from CRAN via `install.packages()`
- Web site: http://ggplot2.org (better documentation)


## What is ggplot2?

- Grammar of graphics represents an abstraction of graphics ideas/objects
- Think “verb”, “noun”, “adjective” for graphics
- Allows for a “theory” of graphics on which to build new graphics and graphics objects
- “Shorten the distance from mind to page”


## Grammer of Graphics

### “In brief, the grammar tells us that a statistical graphic is a __mapping__ from data to __aesthetic__ attributes (colour, shape, size) of __geometric__ objects (points, lines, bars). The plot may also contain statistical transformations of the data and is drawn on a specific coordinate system”

- from _ggplot2_ book

## Plotting Systems in R: Base

- “Artist’s palette” model
- Start with blank canvas and build up from there
- Start with `plot` function (or similar)
- Use annotation functions to add/modify (`text`, `lines`, `points`, `axis`)

## Plotting Systems in R: Base

- Convenient, mirrors how we think of building plots and analyzing data
- Can’t go back once plot has started (i.e. to adjust margins); need to plan in advance
- Difficult to “translate” to others once a new plot has been created (no graphical “language”)
  - Plot is just a series of R commands

## Plotting Systems in R: Lattice

- Plots are created with a single function call (`xyplot`, `bwplot`, etc.)
- Most useful for conditioning types of plots: Looking at how $y$ changes with $x$ across levels of $z$
- Things like margins/spacing set automatically because entire plot is specified at once
- Good for putting many many plots on a screen


## Plotting Systems in R: Lattice

- Sometimes awkward to specify an entire plot in a single function call
- Annotation in plot is not intuitive
- Use of panel functions and subscripts difficult to wield and requires intense preparation
- Cannot “add” to the plot once it’s created


## Plotting Systems in R: ggplot2

- Split the difference between base and lattice
- Automatically deals with spacings, text, titles but also allows you to annotate by “adding”
- Superficial similarity to lattice but generally easier/more intuitive to use
- Default mode makes many choices for you (but you _can_ customize!)


## The Basics: `qplot()`

- Works much like the `plot` function in base graphics system
- Looks for data in a data frame, similar to lattice, or in the parent environment
- Plots are made up of _aesthetics_ (size, shape, color) and _geoms_ (points, lines)


## The Basics: `qplot()`

- Factors are important for indicating subsets of the data (if they are to have different properties); they should be __labeled__
- The `qplot()` hides what goes on underneath, which is okay for most operations
- `ggplot()` is the core function and very flexible for doing things `qplot()` cannot do


## Example Dataset

```{r}
library(ggplot2)
str(mpg) # not the levels/factors
```


## ggplot2 “Hello, world!”

```{r}
qplot(displ, hwy, data = mpg)
```


## Modifying aesthetics

```{r}
qplot(displ, hwy, color = drv, data = mpg)
```


## Adding a geom

```{r}
qplot(displ, hwy, data = mpg, geom = c("point", "smooth"))
```

---

## Histograms

```{r}
qplot(hwy, data = mpg, fill = drv)
```

## Facets

```{r, fig.width=4.5}
qplot(displ, hwy, data = mpg, facets = . ~ drv) # variable_left ~ variable_right => row ~ col
qplot(hwy, data = mpg, facets = drv~., binwidth = 2)
```

## MAACS Cohort

- Mouse Allergen and Asthma Cohort Study
- Baltimore children (aged 5—17)
- Persistent asthma, exacerbation in past year
- Study indoor environment and its relationship with asthma morbidity
- Recent publication: http://goo.gl/WqE9j8

```r
eno <- read.csv("eno.csv")
skin <- read.csv("skin.csv")
env <- read.csv("environmental.csv")
m <- merge(eno, env, by = "id")
maacs <- merge(m, skin, by = "id")
```


## Example: MAACS

```r
str(maacs)
```



## Histogram of eNO

```r
qplot(log(eno), data = maacs)
```

## Histogram by Group

```r
qplot(log(eno), data = maacs, fill = mopos)
```

## Density Smooth

```r
qplot(log(eno), data = maacs, geom = "density")
qplot(log(eno), data = maacs, geom = "density", color = mopos)
```

## Scatterplots: eNO vs. PM$_{2.5}$

```r
qplot(log(pm25), log(eno), data = maacs)
qplot(log(pm25), log(eno), data = maacs, shape = mopos)
qplot(log(pm25), log(eno), data = maacs, color = mopos)
```

## Scatterplots: eNO vs. PM$_{2.5}$

```r
qplot(log(pm25), log(eno), data = maacs, color = mopos, 
      geom = c("point", "smooth"), method = "lm")
```

## Scatterplots: eNO vs. PM$_{2.5}$

```r
qplot(log(pm25), log(eno), data = maacs, geom = c("point", "smooth"), 
      method = "lm", facets = . ~ mopos)
```


## Summary of qplot()

- The `qplot()` function is the analog to `plot()` but with many built-in features
- Syntax somewhere in between base/lattice
- Produces very nice graphics, essentially publication ready (if you like the design)
- Difficult to go against the grain/customize (don’t bother; use full ggplot2 power in that case)




## Basic Components of a ggplot2 Plot
- A _data frame_
- _aesthetic mappings_: how data are mapped to color, size 
- _geoms_: geometric objects like points, lines, shapes. 
- _facets_: for conditional plots.  (multiple panels)
- _stats_: statistical transformations like binning, quantiles, smoothing. 
- _scales_: what scale an aesthetic map uses (example: male = red, female = blue). 
- _coordinate system_ 


## Building Plots with ggplot2
- When building plots in ggplot2 (rather than using qplot) the “artist’s palette” model may be the closest analogy
- Plots are built up in layers
  - Plot the data
  - Overlay a summary
  - Metadata and annotation



## Example: BMI, PM$_{2.5}$, Asthma
- Mouse Allergen and Asthma Cohort Study
- Baltimore children (age 5-17)
- Persistent asthma, exacerbation in past year
- Does BMI (normal vs. overweight) modify the relationship between PM$_{2.5}$ and asthma symptoms?

```r
maacs <- read.csv("ggplot2_lecture2_data.csv")
```

 
## Basic Plot

```r
library(ggplot2)
qplot(logpm25, NocturnalSympt, data = maacs, facets = . ~ bmicat, 
      geom = c("point", "smooth"), method = "lm")
```

 
## Building Up in Layers

```r
head(maacs)
g <- ggplot(maacs, aes(logpm25, NocturnalSympt))
summary(g)
```


 
## No Plot Yet!

```r
g <- ggplot(maacs, aes(logpm25, NocturnalSympt))
print(g)
```



## First Plot with Point Layer

```r
g <- ggplot(maacs, aes(logpm25, NocturnalSympt))
g + geom_point()
```

 
## Adding More Layers: Smooth

```r
g + geom_point() + geom_smooth()
g + geom_point() + geom_smooth(method = "lm")
```

 

## Adding More Layers: Facets

```r
g + geom_point() + facet_grid(. ~ bmicat) + geom_smooth(method = "lm")
```
Make sure that data has proper factors, etc... => will make process easier when plotting
## Annotation
- Labels: `xlab()`, `ylab()`, `labs()`, `ggtitle()`
- Each of the “geom” functions has options to modify 
- For things that only make sense globally, use `theme()` 
  - Example: `theme(legend.position = "none")` 
- Two standard appearance themes are included
  - `theme_gray()`: The default theme (gray background)
  - `theme_bw()`: More stark/plain 


## Modifying Aesthetics

```r
g + geom_point(color = "steelblue", size = 4, alpha = 1/2) # alpha => transparency
g + geom_point(aes(color = bmicat), size = 4, alpha = 1/2) # assigning color to be a value of a variable
```


## Modifying Labels

```r
g + geom_point(aes(color = bmicat)) + labs(title = "MAACS Cohort") + 
  labs(x = expression("log " * PM[2.5]), y = "Nocturnal Symptoms")
```


## Customizing the Smooth

```r
g + geom_point(aes(color = bmicat), size = 2, alpha = 1/2) + 
  geom_smooth(size = 4, linetype = 3, method = "lm", se = FALSE)
```



## Changing the Theme

```r
g + geom_point(aes(color = bmicat)) + theme_bw(base_family = "Times")
```



## A Note about Axis Limits

```{r, fig.width=4.5, fig.height=3}
testdat <- data.frame(x = 1:100, y = rnorm(100))
testdat[50,2] <- 100  ## Outlier!
plot(testdat$x, testdat$y, type = "l", ylim = c(-3,3))
g <- ggplot(testdat, aes(x = x, y = y))
g + geom_line()
```

---

## Axis Limits

```{r, fig.width=4.5}
g + geom_line() + ylim(-3, 3) # outlier is excluded (not good)
g + geom_line() + coord_cartesian(ylim = c(-3, 3)) # use additional option COORD_CARTESIAN
```


## More Complex Example
- How does the relationship between PM$_{2.5}$ and nocturnal symptoms vary by BMI and NO$_2$?
- Unlike our previous BMI variable, NO$_2$ is continuous
- We need to make NO$_2$ categorical so we can condition on it in the plotting
- Use the `cut()` function for this


## Making NO$_2$ Tertiles

```r
## Calculate the tertiles of the data
cutpoints <- quantile(maacs$logno2_new, seq(0, 1, length = 4), na.rm = TRUE)
## Cut the data at the tertiles and create a new factor variable
maacs$no2tert <- cut(maacs$logno2_new, cutpoints)
## See the levels of the newly created factor variable
levels(maacs$no2tert)
```
 

## Final Plot

```r
## Setup ggplot with data frame
g <- ggplot(maacs, aes(logpm25, NocturnalSympt))
## Add layers
g + geom_point(alpha = 1/3) + 
  facet_wrap(bmicat ~ no2tert, nrow = 2, ncol = 4) + 
  geom_smooth(method="lm", se=FALSE, col="steelblue") + 
  theme_bw(base_family = "Avenir", base_size = 10) + 
  labs(x = expression("log " * PM[2.5])) + 
  labs(y = "Nocturnal Symptoms") + 
  labs(title = "MAACS Cohort")
```

 

## Code for Final Plot

```r
## Setup ggplot with data frame
g <- ggplot(maacs, aes(logpm25, NocturnalSympt))
## Add layers
g + geom_point(alpha = 1/3) + 
  facet_wrap(bmicat ~ no2tert, nrow = 2, ncol = 4) + 
  geom_smooth(method="lm", se=FALSE, col="steelblue") + 
  theme_bw(base_family = "Avenir", base_size = 10) + 
  labs(x = expression("log " * PM[2.5])) + 
  labs(y = "Nocturnal Symptoms") + 
  labs(title = "MAACS Cohort")
```
## Summary
- ggplot2 is very powerful and flexible if you learn the “grammar” and the various elements that can be tuned/modified
- Many more types of plots can be made; explore and mess around with the package (references mentioned in Part 1 are useful)










