---
title: "R Programming (Coursera)"
output: html_notebook
---

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

```{r}
plot(cars)
```

## Vectors and Lists
### Basic Idea
```{r}
x <- c(0.5, 0.6) # created a numerical vector of length 2
# c => concatenate 
x <- c(T, F)
x <- c('a', 'b', 'c')
x <- c(1i, 2i )
```

### Mixing Objects (coercion allows every element in the vector to be the same class)

```{r}
y <- c(1.4, 'a')
y <- c(T, 1, 2)
y <- c(T, "hello")
```

### Explicit Coercion
```{r}

x <- 0:6
class(x)
as.numeric(x)
as.logical(x)
as.character(x)

```
### Some times Coercion doesn't work
```{r}
x <- c('a', 'b', 'c')
as.numeric(x)
as.logical(x)
```

## Lists
```{r}
x <- list(1, "a", TRUE, 1 + 4i)
x # indexed by double brackets
```

# Matrices - special type of vector in R (with dimension)
```{r}
m <- matrix(nrow = 2, ncol = 3)
m
dim(m) # dimension fst = row, snd = col

attributes(m)
```
## Constructed column-wise (not row-wise)
```{r}
m <- matrix(1:6, nrow = 2, ncol = 3)
m
```
# Create matrix differently
```{r}
m <- 1:10 
m 
dim(m) <- c(2,5) # assign the dimensions of the vector m => matrix m
m
```
## CBIND and RBIND
```{r}
x <- 1:3
y <- 10:12
cbind(x,y) # each vector is a column
rbind(x,y) # each vector is a row
```
# Factors (represent categorical data)
```{r}
x <- factor(c("yes", "no", "no", "yes", "yes"))
x

table(x) # frequency count

unclass(x) # how they are represented by the R system

```
## Factors Order of levels
```{r}
x <- factor(c("yes", "yes", "yes", "no", "yes", "yes", "yes"))
x # order of level is determined by alphabetical order (i.e. no is base level)

# can specify with LEVELS
x <- factor(c("yes", "yes", "yes", "no", "yes", "yes", "yes"), levels = c("yes", "no"))
x
```
## Missing Values (NA or NaN)
NaN is used for undefined mathematical operations and Na is used for everything else. 
Na can also have class
NaN is also a Na BUT reverse is not true
```{r}
x <- c(1,2,3,4,NA, 10)
is.na(x)
is.nan(x)
x <- c(1,NaN,3,4,NA, 10)
is.na(x)
is.nan(x)
```
# Data Frame (to store tabular data)
special type of list where every element of the list has to have the same len. 
special attributes => row.names
created => read.table() or read.csv()
can be converted to matrix => data.matrix() (i.e. expect coercion)
```{r}
x <- data.frame(foo = 1:4, bar = c(T, T, F, F)) # each is a column 
x
ncol(x)
nrow(x)

```
## Names 
```{r}
x <- 1:3
names(x) # default is no name
names(x) <- "foo" # only first elt has name
x 
names(x) <- c("foo", "bar", "norf") # each element has name
x 
names(x) # list names

```

```{r}
x <- list(a = 1, b = 2, c = 3)
x
```

```{r}
m <- matrix(1:4, nrow = 2, ncol = 2)
dimnames(m) <- list(c("a", "b"), c("c", "d")) # assign as list of 2 vectors: row names, col names
m
```

# Reading in Large Datasets with read.table / read.csv
* Make rough calc of memory required to store data. If dataset is greater than the amount of RAM available, stop! (# of elements x 8 bytes/numeric => divide by 2^20 bytes/MB => multiply by two to account for overhead required)
* set comment.char = "" if no commented lines exist in code (makes it faster!)
* use the colClasses argument (to make more efficient) - read first couple rows and define it for whole dataset
```{r}
#initial <- read.table("some_data-base.txt", comment.char = "", nrows = 100) # read 100 lines
#classes <- sapply(initial, class) # determine classes of each coloumn
#tabAll <- read.table("some_data-base.txt", comment.char = "", nrows = 100, colClasses = classes)
```
* set nrows (can overestimate)



# Textual Data Formats (which contain more metadata)
* ``dumping`` and ``dputing`` are useful b/c they preserve the metadata

```{r}
y <- data.frame(a = 1, b = "a") # coloumn name included

dput(y) # metadata included
dput(y, file = "y.R") # make file that allows us to  re-construct the object
new.y <- dget("y.R")
new.y

```

Multiple objects can be deparsed using the dump function and read back in using source
```{r}
x <- "foo"
y <- data.frame(a = 1, b = "a") 
dump(c("x", "y"), file = "data.R") # arg -> names of the objects we want to store and the file name in which they should be stored
rm(x,y) # remove them
source("data.R")
y 
x
```

# Interfaces to the Outside World 
* ``file`` open connection to a file
* ``gzfile`` open connection to a gzip compressed file
* ``bzfile`` open connection to a bzip2 compressed file
* ``url`` open connection to a webpage

```{r}
con <- file("y.R", "r")
#data <- read.csv(con)
close(con)
```


```{r}
#con <- gzfile("somefile.gz", "r")
x <- readLines(con, 10) # read 10 lines
x
close(con)

```

```{r}
con <- url("http://www.jhsph.edu", "r")
x <- readLines(con, 10)
head(x)
```




# SUBSETTING
[ always returns object of same class as original (e.g. subset a list with [] -> list) -> can be used to select more than one element (one exception case, however)

[[ used to extract elts of a list or data frame, can only be used to extract a single element and the class of the returned object will not necc. be a list or a df

$ used to extract elements of a list or df BY NAME; semantics are similiar to that of [[

```{r}
x <- c("a", "b", "c", "d") # character vector
x[1] # returns a character vector
x[2] # returns a character vector
x[1:4] # returns a character vector (using numeric index)
x[x>"a"] # filtering for letters that are greater than 'a'
u <- x > "a" # logical vector u that tells us which elts of x are greater than 'a'
u
x[u]
```


## Subsetting Lists
```{r}
x <- list(foo = 1:4, bar = 0.5) # list consisting of two names elements ... whose value is ...
x[1] # returns a list  (same class as original)
x[[1]] # a sequence
x$bar # returns element associated with name bar
x[["bar"]] # same operation as above
x["bar"] # returns list


# to subset multiple elements of a list...
x <- list(foo = 1:4, bar = 0.3, baz = "hello")
x[c(1,3)]
```

```{r}
x <- list(foo = 1:4, bar = 0.3, baz = "hello")
name <- "foo"
x[[name]] # computed index for 'foo' - finds elt associated with 'foo' not 'name'
x$name # dollar sign is weaker as it literally searches for elt associated with 'name'
x$foo # this is correct use of $
```

Subsetting Nested Elements of a List
```{r}
# [[ can take an integer sequence (nested/repeated subsetting)
x <- list(a = list(10, 12, 14), b = c(3.21, 2.32))
x[[c(1,3)]] # same as double subsetting
x[[1]][[3]] # equivalent to above
x[[c(2,1)]]

```

## Subsetting Matrices
```{r}
x <- matrix(1:6, 2, 3)
x
x[1,2]
x[2,1]
```
```{r}
x[1,] # first row
x[,2] # second col
```

```{r}
x <- matrix(1:6, 2, 3)
x[1,2] # default behavior is to return a vector of length 1, instead of 1x1 matrix

x[1,2,drop = FALSE] # don't DROP the dimension => returns 1x1 matrix

```

```{r}
x <- matrix(1:6, 2, 3)
x[1,] # again, drop = TRUE as default and dimensions are dropped => vector returned
x[1, , drop = FALSE] 

```

## Subsetting - Partial Matching
partial matching of names is allowed with [[ and $.
```{r}
x <- list(aardvark = 1:5)
x$a # as only name that starts with a is aardvark, R understands this and runs x$aardvark
x[["a"]] # by default, partial matching is turned off with [[
x[["a", exact = FALSE]] # specify exact as FALSE, it gives what we want as the aardvark element is what matches the closest
```

# Removing NA Values
Common Operation - many missing values 
key: Create logical vector that tells us where the NAs are such that we can remove them by subsetting
```{r}
x <- c(1,2,3,4,5,NA, 4,4,5,6,NA, 299, 21)
bad <- is.na(x)
bad # false => not NA, true => NA (i.e. BAD Data)
x[!bad] # ! => called bang operator (i.e. bang bad => good) similar to not
```
When considering multiple objects at once
```{r}
x <- c(1,2,3,4,5,NA, 4,4,5,6,NA, 299, 21)
y <- c('a','ac','s','s',' ', NA, 'a', 'bc', 'e', 'q',NA, 299, '21') # note the coercion
good <- complete.cases(x,y) # use on both vectors => which positions/indices contain non-NA values for both objects/vectors
good

x[good]
y[good]
```

```{r} 
airquality[1:6, ] # some missing values... we only want rows in df where all values are present (non-missing)
good <- complete.cases(airquality) # logical vector which tells us which rows are complete
head(good)

airquality[good, ][1:6, ] # of the good rows, show the first 6 and all cols
```

# Vectorized Operations
Benefit for command line use 
More efficient, concise, and easier to read
```{r}
x <- 1:4; y <-6:9
# to add the two vectors... (add in parellel) no need to loop! simply...
x + y
# Logical vectors
x>2 # compares all the numbers... 
x>=2
y==8 # equality
x*y
x/y
```
Similarly, in matrices...
```{r}
x <- matrix(1:4, 2,2) ; y <- matrix(rep(10,4), 2,2) # rep => 10, 10, 10, 10
# ELEMENT-WISE MULTIPLICATION
x*y
x/y
# TRUE MATRIX MULTIPLICATION
x %*% y

```

## Practice Questions
```{r}
hw1_data <- read.csv("~/Xenograft_RStudio/hw1_data.csv")
#View(hw1_data)
bad <- is.na(hw1_data[,1])
mean(hw1_data[,1][!bad])

ourfilter <-   hw1_data$Ozone>31 & hw1_data$Temp>90 & complete.cases(hw1_data$Ozone, hw1_data$Temp)
mean(hw1_data[,2][ourfilter])

mean(hw1_data$Temp[hw1_data$Month==6])

max(hw1_data$Ozone[!is.na(hw1_data$Ozone)][hw1_data$Month==5])
```

# Notes from SWIRL
| Now, let's create a small collection of numbers called a vector. Any object that contains data is called
| a data structure and numeric vectors are the simplest type of data structure in R. In fact, even a single
| number is considered a vector of length one.
| The easiest way to create a vector is with the c() function, which stands for 'concatenate' or 'combine'.
| To create a vector containing the numbers 1.1, 9, and 3.14, type c(1.1, 9, 3.14).
| Type z to view its contents. Notice that there are no commas separating the values in the output.


| When we did z * 2 + 100 in our earlier example, z was a vector of length 3, but technically 2 and 100 are
| each vectors of length 1.
| Behind the scenes, R is 'recycling' the 2 to make a vector of 2s and the 100 to make a vector of 100s. In
| other words, when you ask R to compute z * 2 + 100, what it really computes is this: z * c(2, 2, 2) +
| c(100, 100, 100).

| Vectors come in two different flavors: atomic vectors and lists. An atomic vector contains exactly one data type, whereas a list may contain multiple data types. We'll explore atomic vectors further before we get to
| lists.


| Index vectors come in four different flavors -- logical vectors, vectors of positive integers, vectors of negative integers, and vectors of character strings -- each of which we'll cover in this lesson.

| Many programming languages use what's called 'zero-based indexing', which means that the first element of a vector is considered element 0. R uses 'one-based indexing', which (you guessed it!) means the first element of a
| vector is considered element 1.

| The main difference, as you'll see, is that matrices can only contain a single class of data, while data frames can consist of many different classes of data. (kind of like vectors vs lists)
| The example that we've used so far was meant to illustrate the point that a matrix is simply an atomic vector with a dimension attribute. A more direct method of creating the same matrix uses the matrix() function.
| This is called 'implicit coercion', because we didn't ask for it. It just happened. But why didn't R just convert the names of our patients to numbers? I'll let you ponder that question on your own.



# Week 2 - Control Structures
* if, else: testing a condition
* for: execute a loop a fixed number of times
* while: execute a look while condition is true
* repeat: execute infinite loop
* break: break the execution of loop
* next: skip an iteration of a loop (i.e. continue in python)
* return: exit a function

## If-else
Else branch is optional (i.e. if(){} if(){} is possible)
```{r}
x <-4

if(x>3){
    y <- 10
} else {
    y <- 0
}

# or same process 

y <- if(x>3){
    10
} else {
    0
}
```

## For-loop
```{r}
for(i in 1:10){
    print(i)
}

x<-c('a', 'b', 'c', 'd')
for(i in 1:4){
    print(x[i]) # prints the i-th element of vector x
}

# using seq_along, we don't need to know the length of x
for(i in seq_along(x)){
    print(x[i])
}

for(letter in x){
    print(letter)
}

for(i in 1:4) print(x[i]) # if one line, can omit the curly brackets => not best practice


```
Nested for-loops
```{r}
x<-matrix(1:6,2,3)
for(i in seq_len(nrow(x))){ # seq_len produces sequence based on length -> i.e. based on number of rows -> 1:2
    for(j in seq_len(ncol(x))){
        print(x[i,j]) # print every element of the matrix (row priority/order)
    }
}
```
Be careful with repeated nesting - complex != good (but simple = good)

## While-loop
```{r}

count <- 0
while(count<10){
    print(count)
    count <- count + 1
}
```
Be careful to ensure it doesn't become an infinite loop 

```{r}
z <- 5
while (z>=3 && z<=10){
    print(z)
    coin <- rbinom(1,1,0.5)
    
    if(coin ==1){
        z <- z+1
    } else {
        z <- z-1
    }
}
```
Conditions are always evaluated from left to right (when given multiple condition operations)

## Repeat, Next, Break
Repeat - requires break
```{r}
x0 <- 1
tol <- 1e-8

#repeat {
#  x1<- computeEstimate()
#  
#  if(abs(x1-x0)<tol){
#    break
#  } else {
#    x0 <- x1
#  }
#}
```
Not the best idea - dangerous. 

## next, return
```{r}
for(i in 1:10){ # this loop does nothing
  if(i<=20){
    next
  }
}
```

Generally speaking, infinite loops are not reccomended even if they are theoretically correct


## Simple R Functions
```{r}
add2 <- function(x,y){
    x+y
}

above10 <- function(x){
    use <- x>10 # logical vector
    x[use]
}

above <- function(x, n =10){ # default to 10
    use <- x>n # logical vector
    x[use]
}
```

```{r}
# mean of each column

columnmean_error <- function(y){
    nc <- ncol(y)
    means <- numeric(nc) # numeric vector of length nc - initialized as all 0s
    for(i in 1:nc){
        means[i] <- mean(y[,i])
    }
    means
}

columnmean <- function(y, removeNA = TRUE){
    nc <- ncol(y)
    means <- numeric(nc) # numeric vector of length nc - initialized as all 0s
    for(i in 1:nc){
        means[i] <- mean(y[,i], na.rm = removeNA)
    }
    means
}

columnmean_error(airquality)
columnmean(airquality)
columnmean(airquality, FALSE)
```


# Functions
functions in R are "first class objects" => treated much like any other R object (i.e. functions can be passed as arg for other functions)

* formal arguments => included in function definition
* the ``formals`` returns list of all formal arguments
```{r}
mydata = rnorm(100)
sd(mydata)
sd(x = mydata) # arg is named
sd(x=mydata, na.rm=FALSE)
sd(na.rm=FALSE, x=mydata)
sd(na.rm=FALSE, mydata) # if unnamed -> follow order
```
arg matching
```{r}
args(lm)
#lm(data = mydata, y-x, model = FALSE, 1:100) # confusing but works (not recommended)
#lm(y-x, mydata, 1:100, model = FALSE)
```

R thought process: absolute match, partial match, positional match of arguments


Lazy Evaluation
```{r}
f <- function(a,b){
  a^2
}
f(2) # No error produced because b isn't used anyways - 2 gets positionally matched to a
```

```{r}
f <- function(a,b){
  print(a)
  print(b)
}
f(45) # execute until error!
```


``...`` indicate a variable number of args that are usually passed on to other functions
```{r}
myplot <- function(x,y,type="l", ...){
  plot(x,y,type = type, ...)
}

# Generic functions
mean

# unknown number of arguments 
args(paste)
args(cat)
```
When using ``...``, any arg appearing after ... must be named explicity (no partial matching)!


# Scoping Rules 
## Symbol Binding
when R tries to bind a value (e.g. my function, etc..) to a symbol (e.g. lm, mean, etc...), it searches through a series of environments to find the appropriate value. First, search: 
1. Search the Global Environment
2. Search the namespace of each packages on the search list
```{r}
search()
```
When user loads package with ``library``, the namespace of that package gets put in position two of the search list.

### SCOPING RULES
LEXICAL/STATIC SCOPING and DYNAMIC SCOPING

```{r}
f <- function(x,y){
  x^2 + y/z    # z => free variable
}
```

how do we assign a value to a free variable? SCOPING RULES

## Lexical Scoping in R: the values of free variables are searched for in the environment in which the function was defined (continue search in parent env) 


environment: a collection of (symbol, value) pairs (e.g. x is symbol, 3.14 is value)
* every environment has parent environment (multiple children is possible)
* a function + an environment = a closure or function closure


## Lexical Scoping - R
* Typically a function is defined in the global environment

In R, the free variable bindings are resolve by first looking in the environment in which the function was created.

```{r}
make.power <- function(n){
  pow <- function(x){ # function defined within a function. 
    x^n
  }
  pow
}

cube <- make.power(3) # pow is lacking its argument (n is defined as 3) => function cube 
square <- make.power(2)

# How do you know what's in a fxn's environment? 
ls(environment(cube)) # to look at environment in which a function is defined. 
# in the environment there is an object called "n" and "pow"
get("n", environment(cube)) # what's the value of "n" in the environment of cube (this fxn)

ls(environment(square))
get("n", environment(square))
```

Lexical vs. Dynamic Scoping
```{r}
y <- 10

f <- function(x){
    y<-2
    y^2 + g(x)
}

g <- function(x){ 
    x*y}

#what is the value of 
f(3)
# lexical -> value of y in g is 10 (searched in calling environment / environment where function is called)
# dynamic -> y in g is 2

```


```{r}
g <- function(x){
  a<- 3
  x+a+y
}

g(2)

y <-3
g(2)
```
Lexical Scoping is supported in: 
* Scheme
* Perl
* Python
* Common Lisp (all languages converge to Lisp)

### Consequences of Lexical Scoping 
* In R, all objects must be stored in memory 
* All functions must carry a pointer to their defining environment, which could be anywhere
* In S-PLUS, free variables are always looked up in the global workspace

parent frame => environment in which a function was called


Why does this all matter? 
## Optimization
Easier to write cleaner, more organized code in solving optimization problems. 
```{r}
# write a constructor function
make.NegLogLik <- function(data, fixed=c(FALSE, FALSE)){ 
    params <- fixed
    function(p) {
        params[!fixed] <- p 
        mu <-params[1]
        sigma <- params[2]
        a <- -0.5*length(data)*log(2*pi*sigma^2)
        b <- -0.5*sum((data-mu)^2) / (sigma^2)
        -(a+b)
    }
}

set.seed(1); normals <- rnorm(100,1,2)
nLL <- make.NegLogLik(normals)
nLL
ls(environment(nLL))
#get("data", environment(nLL))

optim(c(mu=0,sigma=1), nLL)$par


#fixing sigma = 2
nLL <- make.NegLogLik(normals, c(FALSE,2))
optimize(nLL, c(-1,-3))$minimum

#fixing mu = 1
nLL <- make.NegLogLik(normals, c(1,FALSE))
optimize(nLL, c(1e-6,10))$minimum

```
```{r}
nLL <- make.NegLogLik(normals, c(1,FALSE))
x <- seq(1.7, 1.9, len =100)
y <- sapply(x,nLL)
plot(x, exp(-(y-min(y))), type='l')

nLL <- make.NegLogLik(normals, c(FALSE, 2))
x <- seq(0.5, 1.5, len =100)
y <- sapply(x,nLL)
plot(x, exp(-(y-min(y))), type='l')
```

# Coding Standards

1. use text editor/files (ASCII text)
2. Indent your code
3. Limit the width of your code (80 columns?)
4. Limit the length of individual functions (one purpose/function)


## Dates and Times in R
* Dates are represented by ``Date`` class
* Times by ``POSIXct`` and ``POSIXlt``

``weekdays``
```{r}
x <- as.Date("1970-01-01")
x
unclass(x) # days since 1970-01-01 is how its stored internally

```
```{r}
x <- Sys.time()
x
p <- as.POSIXlt(x)
names(unclass(p))
p$sec
```
```{r}
datestring <- c("January 10, 2012 10:40", "December 9, 2011 9:10")
x <- strptime(datestring, "%B %d, %Y %H:%M")
x

class(x)

x <- as.Date("2012-01-01")
y <- strptime("9 Jan 2011, 10:40:29", "%d %b %Y %H:%M:%S")
#-y gives error
x <- as.POSIXlt(x)
x-y

```

```{r}
x <- as.Date("2012-03-01")
y <- as.Date("2012-02-28")
x -y
x <- as.POSIXct("2012-10-25 01:00:00")
y <- as.POSIXct("2012-10-25 06:00:00", tz = "GMT")
y-x
```


# Loop Functions
##lapply
arg) a list, a function (or name of it), other arguments via its .. argument

```{r}
(lapply)
```
```{r}
# x is a list of two vectors/elements
x <- list(a=1:5, b = rnorm(10)) # what goes in may/may not be a list => output is always a list
lapply(x, mean)
```

```{r}
x <- list(a = 1:4, b = rnorm(10), c = rnorm(20,1), d = rnorm(100, 5))
lapply(x, mean) # apply mean on every element of the list
```

```{r}
x <- 1:4
lapply(x, runif) # runif => random uniform generator 

lapply(x, runif, min = 0, max = 10)
```

lapply and friends make heavy use of anonymous functions 
```{r}
x <- list(a = matrix(1:4, 2,2), b = matrix(1:6, 3,2))

lapply(x, function(elt) elt[,1]) # get first column of a and b
```
``sapply``: simplifies result of lapply. 
* if result is a list where every element is length 1 => return vector
* if result is a list where every element has same length => return matrix
* else, function the same way as lapply (i.e. return a list)
```{r}
x <- list(a = 1:4, b = rnorm(10), c = rnorm(20,1), d = rnorm(100, 5))
sapply(x, mean)
```

## apply
* used to apply a function to the rows or columns of a matrix
* it can be used with general arrays
* it is not really faster than writing a loop (but it works in one line)

```{r}
str(apply) # ... are args that our function may require

x <- matrix(rnorm(200), 20, 10)
apply(x, 2, mean) # keep 2nd dimension and collapse the 1st dimension (rows) - preserve col, collapse the rows
print("Or...")
apply(x, 1, mean) # preserve the rows and collapse the col (i.e. 2nd dimension)

```
These following are optimized
* ``rowSums``
* ``rowMeans``
* ``colSums``
* ``colMeans``
These are much faster

```{r}
x <- matrix(rnorm(200), 20, 10)
apply(x, 1, quantile, probs = c(0.25,0.75)) # for each row, the quantiles are computed and displayed
```

```{r}
a <- array(rnorm(2*2*10), c(2,2,10))# average of the two by two matrices
apply(a, c(1,2), mean) # preserve the 1st and 2nd dimension (row/col) and collapse the 3rd dimension using the function
rowMeans(a, dims = 2) # alternatively, dims = 2 does the same trick
```

## mapply
* multivariate apply of sorts which applies a function in parellel over a set of arguments
```{r}
str(mapply)
```
... => arguments to apply over (objects)
```{r}
list(rep(1,4), rep(2,3), rep(3,2), rep(4,1))

# instead we can do 
mapply(rep, 1:4, 4:1)

```
Instant Vectorization
```{r}
noise <- function(n, mean, sd) {
  rnorm(n, mean, sd)
}
noise(5,1,2)

mapply(noise, 1:5, 1:5, 2)
# same as...
list(noise(1,1,2), noise(2,2,2), noise(3,3,2), noise(4,4,2), noise(5,5,2))
```


## tapply
apply a function over subsets of a vector 
```{r}
str(tapply) 
```
INDEX: factor or a list of factors (or else they are coerced to factors)
```{r}
x <- c(rnorm(10), runif(10), rnorm(10,1))
factor <- gl(3, 10)
factor

tapply(x,factor,mean)
tapply(x,factor,mean, simplify = FALSE)
tapply(x, factor, range)

```
## Split
* the first step of tapply => function that takes vector or other objects and splits it into groups determined by a factor or a list of factors
```{r}
str(split)
```
```{r}
x <- c(rnorm(10), runif(10), rnorm(10,1))
factor <- gl(3, 10)
split(x, factor)
```

```{r}
sapply(split(x, factor), mean)
tapply(x, factor, mean)
```
```{r}
library(datasets)
head(airquality)
```

```{r}
# mean within each month
s <- split(airquality, airquality$Month)
lapply(s, function(x) colMeans(x[, c("Ozone", "Solar.R", "Wind")]))
```

```{r}
sapply(s, function(x) colMeans(x[, c("Ozone", "Solar.R", "Wind")]))
sapply(s, function(x) colMeans(x[, c("Ozone", "Solar.R", "Wind")], na.rm = TRUE))

```
Splitting on More than one level
```{r}
x <- rnorm(10)
f1 <- gl(2,5)
f2 <- gl(5,2)
f1
f2
interaction(f1,f2)
```

```{r}
str(split(x, list(f1,f2))) # interaction function is implicitly called
# some levels are empty... 

str(split(x, list(f1,f2), drop = TRUE)) 
```


# Debugging
## Questions to ask: 
* What was your input? How did you call the function? 
* What were you expecting? Output, messages, other results? 
* What did you get? 
* How does what you get differ from what you were expecting? 
* Were your expectations correct in the first place? 
* Can you reproduce the problem (exactly)? 


## Basic Tools for Debugging 
* ``traceback``: prints out the function call stack after an error occurs. 
* ``debug``: flags a function for 'debug' mode which allows you to step through execution of a function one line at a time
* ``browser``: similar to debug fxn. suspends the execution of a function wherever it is called
* ``trace``: allows you to insert debugging code into a function a specific places
* ``recover``: allows you to modify the error behavior s.t. you can browse the function call stack


## Traceback 
```{r}
mean(u)
traceback() # only immediately afterwards
```
```{r}
lm(u ~ o)
traceback()
```
```{r}
debug(lm)
lm(u~o)
# press n and enter in the console
```

## recover
```{r}
options(error = recover)
read.csv("filethatdoesn'texist") # select number to look at the environment 
```

| Try vapply(flags, unique, numeric(1)), which says that you expect each element of the result to be a numeric vector of
| length 1. Since this is NOT actually the case, YOU WILL GET AN ERROR. Once you get the error, type ok() to continue to the
| next question.
| Use tapply(flags$animate,
| flags$landmass, mean) to apply the mean function to the 'animate' variable separately for each of the six landmass groups,
| thus giving us the proportion of flags containing an animate image WITHIN each landmass group.



# Str Function
* a diagnostic functoin and an alternative to 'summary'
* suited to compactly display the contents of possibly nested lists
* what's in this object? 

```{r}
str(str)
```

```{r}
str(lm)
str(ls)
```

```{r}
x <- rnorm(100, 2, 4)
summary(x)
str(x)

f <- gl(40, 10)
str(f)
summary(f)
```

```{r}
str(airquality)
```
Good for quick examination
```{r}
m <- matrix(rnorm(100), 10, 10)
str(m) # first columns
m[,1]
```

for lists...
```{r}
s <- split(airquality, airquality$Month)
str(s)
```

# Simulations
* rnorm: generate random normal variates within given mean and sd
* dnorm: evaluate the normal probability density at a point
* pnorm: evaluate the cumulative distribution functin for a normal distribution
* rpois: generate random Poisson variates with a given rate

Probability Distribution functions. PREFIXES are..
* d for density
* r for random number generation* p for cumulative distribution
* q for quantile function
```{r}
str(dnorm)
str(qnorm)
str(pnorm)
str(rnorm)
```
lower.tail = FALSE => upper.tail

if $\phi$ is the cumulative distribution dunction for a standard Normal Distribution, then pnorm(q) = $\phi(q)$ and qnorm(p) = $\phi^{-1}(p)$.

```{r}
x <- rnorm(10)
x
x <- rnorm(10, 20, 2)
```

Set the seed!
```{r}
set.seed(1)
rnorm(5)
rnorm(5)
set.seed(1)
rnorm(10)
```

```{r}
rpois(10,1) # 10 random generated with rate = 1
ppois(2,2) # Pr(x <=2)
ppois(4,2)

```


## Simulate a Linear Model 
```{r}
set.seed(2)
x <- rnorm(100) 
e <- rnorm(100, 0, 2)
y <- 0.5 + 2 * x + e
summary(y)
plot(x,y)
```
```{r}
set.seed(10)
x <- rbinom(100, 1, 0.5) 
e <- rnorm(100, 0, 2)
y <- 0.5 + 2 * x + e
summary(y)
plot(x,y)
```
 
```{r}
set.seed(1)
x <- rnorm(100) 
log.mu <- 0.5 + 0.3*x
y <- rpois(100, exp(log.mu))
summary(y)
plot(x,y)
```

## Sample Function
draws randomly from a specified set of scalar objects allowing you to sample from arb. distributions
```{r}
set.seed(1)
sample(1:10, 4)
sample(1:10, 4)
sample(letters, 5)
sample(1:10) ## Permutation
sample(1:10) ## Permutation
sample(1:10, replace = TRUE)
```


# R Profiler
"Premature optimization is the root of all evil
- Design FIRST, then OPTIMIZE"

```{r}
system.time(readLines("http://www.jhsph.edu"))
```
```{r}
hilbert <- function(n) { 
  i <- 1:n
  1/outer(i -1, i, "+")}

x <- hilbert(1000) 
system.time(svd(x))
```


| For categorical variables (called 'factor' variables in R), summary() displays the number of times each value (or 'level')
| occurs in the data. For example, each value of Scientific_Name only appears once, since it is unique to a specific plant. In
| contrast, the summary for Duration (also a factor variable) tells us that our dataset contains 3031 Perennial plants, 682
| Annual plants, etc.

























